### 2018-09-03 paper/casestudies.tex
added discussion + more suggestions

For implementing the test case $U_F(p)$ with sub-processes $U_F(p,s)$, it is
advisable to avoid an enumeration of traces $s$ of the reference process.
Instead, we calculate the following auxiliary functions from $P$'s transition
graph.
%
\begin{eqnarray*}
\text{\it initials} & : & N \fun \power (\Sigma)
\\
\minhits & : & N \fun \power\power(\Sigma)
\end{eqnarray*}
%
In a state $n = G(P)/s$, the set $\text{\it initials}(n)$ equals the events
labelling outgoing edges of $n$, so  $\text{\it initials}(n) = [P/s]^0$. Function
$\minhits$ maps $n$ to the set of all minimal hitting sets associated
with the minimal acceptances of $n$, so $\minhits(n) =
\minhits(P/s)$. Using the transition function $t$ and the above
functions, $U_F(p)$ can be re-written as the failures-equivalent CSP process
below.
%
\begin{eqnarray}
U_F^1(p) & = & U_F^1(p,0,\ii n)
\\
U_F^1(p,k,n) & = & \big(\Extchoice e:(\Sigma - \text{\it initials}(n)) @ e \then \efail\then \Stop \big)
\label{eq:xufa}
\\ & & \extchoice \nonumber
\\ & & (\minhits(n) = \varnothing)    \&   \big( \epass \then \Stop \big)
\label{eq:xufb}
\\ & & \extchoice \nonumber
\\ & & (k < p) \& \big(\Extchoice e:\text{\it initials}(n) @ e \then U_F^1(p,(k+1),t(n,e)) \big)
\label{eq:xufc}
\\ & & \extchoice \nonumber
\\ & & (k = p) \& \big( \sqcap_{H\in\minhits(n)} (\Extchoice e:H @ e \then \epass \then\Stop ) \big)
\label{eq:xufd}
\end{eqnarray}

@@

\begin{example}
\label{ex:uf1tests} Consider the following erroneous implementation $Z$ of
process $P$ from Example~\ref{ex:a} from the point of view of failures
refinement.
\begin{eqnarray*}
Z & = & a \then (Q_1 \intchoice R_1(r_{max},0))
\\
Q_1 & = & a\then Z \extchoice c\then Z
\\
R_1(r_{max},k) & = & (k < r_{max}) \& \big( c\then R_1(r_{max},k+1)\extchoice b\then Z  \big)
\\ & & \extchoice
\\ & & (k = r_{max}) \& \big( c\then R_1(r_{max},r_{max})\intchoice b\then Z  \big)
\end{eqnarray*}
It is easy to see (and can be checked with FDR) that $Z$ is trace-equivalent
to $P$. While $k < r_{max}$, $Z$ also accepts the same sets of events as $P$.
When $R_1(r_{max},k)$ runs through several recursions and $k = r_{max}$ is
fulfilled, however, $R_1(r_{max},k)$ makes an internal choice, instead of
offering an external choice, and refinement does not hold.
% alcc: I don't think there is a notion of refinement for state?
%so this process state no
%longer refines the corresponding process state $R$ of the reference process
%$P$ in the failures model.
Fig.~\ref{fig:tgZ} shows the normalised transition graph of $Z$ for $r_{max}
= 3$.

Running the test $U_F^1(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p =
4$ states and $G(Z)$ has $q=5$, so $pq=20$ is an upper bound for the test
depth to be used according to Theorem~\ref{th:failurestest}), tests
$U_F^1(0),\dots, U_F^1(3)$ are passed by $Z$, but $Z$ fails $U_F^1(4)$,
because after  execution trace
\[
s = a.c.c.c, \qquad\text{(note that $G(Z)/s = \text{Node\_3}$ according to Fig.~\ref{fig:tgZ})},
\]
it may be the case that $Z$ accepts only $\{b\}$ due to the internal choice
and $U_F^1(4)$ -- also due to internal choice -- accepts only the minimal
hitting set $\{ c \}$ and the event $a\in (\Sigma - [P/s]^0)$. So,
$(Z\parallel[\Sigma] U_F^1(4))/s$ deadlocks, and the $pass$ event cannot be
produced. Another failing execution arises if $Z/s$ chooses to accept only
$\{c \}$, while $U_F^1(4)/s$ choses to accept only $\{a,b\}$. Therefore,
\[
(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F^1(4)),
\]
and the test fails.
\xbox
\end{example}

>>>>

For implementing the test case $U_F(p)$ with sub-processes $U_F(p,s)$, it is
advisable to avoid an enumeration of traces $s$ of the reference process.
Instead, we calculate the following auxiliary functions from $P$'s transition
graph.
%
\begin{eqnarray*}
\text{\it initials} & : & N \fun \power (\Sigma)
\\
\minhits & : & N \fun \power\power(\Sigma)
\end{eqnarray*}
%
As mentioned, in a state $n = G(P)/s$, the set $\text{\it initials}(n)$
equals the events labelling outgoing edges of $n$, so  $\text{\it
initials}(n) = [P/s]^0$. The function $\minhits$ maps $n$ to the set of all
minimal hitting sets associated with the minimal acceptances of $n$, so
$\minhits(n) = \minhits(P/s)$. Using the transition function $t$ and the
above functions, $U_F(p)$ can be re-written as the failures-equivalent CSP
process below.
%
\begin{eqnarray}
U_F^1(p) & = & U_F^1(p,0,\ii n)
\\
U_F^1(p,k,n) & = & \big(\Extchoice e:(\Sigma - \text{\it initials}(n)) @ e \then \efail\then \Stop \big)
\label{eq:xufa}
\\ & & \extchoice \nonumber
\\ & & (\minhits(n) = \varnothing) \& \big( \epass \then \Stop \big)
\label{eq:xufb}
\\ & & \extchoice \nonumber
\\ & & (k < p) \& \big(\Extchoice e:\text{\it initials}(n) @ e \then U_F^1(p,(k+1),t(n,e)) \big)
\label{eq:xufc}
\\ & & \extchoice \nonumber
\\ & & (k = p) \& \big( \sqcap_{H\in\minhits(n)} (\Extchoice e:H @ e \then \epass \then\Stop ) \big)
\label{eq:xufd}
\end{eqnarray}
\fixme{alcc: I think the example should be for the theory presented before,
not with a different notion of test. In particular, 33 is very different.}

@@

\begin{example}
\label{ex:uf1tests} Consider the following implementation $Z$ of process $P$
from Example~\ref{example:CSP} that is erroneous from the point of view of
failures refinement.
\begin{eqnarray*}
Z & = & a \then (Q_1 \intchoice R_1(r_{max},0))
\\
Q_1 & = & a\then Z \extchoice c\then Z
\\
R_1(r_{max},k) & = & (k < r_{max}) \& \big(b\then Z \extchoice  c\then R_1(r_{max},k+1)\big)
\\ & & \extchoice
\\ & & (k = r_{max}) \& \big(b\then Z \intchoice c\then R_1(r_{max},r_{max})\big)
\end{eqnarray*}
It is easy to see (and can be checked with FDR) that $Z$ is trace-equivalent
to $P$. While $k < r_{max}$, $Z$ also accepts the same sets of events as $P$.
When $R_1(r_{max},k)$ runs through several recursions and $k = r_{max}$ is
fulfilled, however, $R_1(r_{max},k)$ makes an internal choice, instead of
offering an external choice, and refinement does not hold.
% alcc: I don't think there is a notion of refinement for state?
%so this process state no
%longer refines the corresponding process state $R$ of the reference process
%$P$ in the failures model.
Fig.~\ref{fig:tgZ} shows the normalised transition graph of $Z$ for $r_{max}
= 3$.

Running the test $U_F^1(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p =
4$ states and $G(Z)$ has $q=5$, so $pq=20$ is an upper bound for the test
depth to be used according to Theorem~\ref{th:failurestest}), tests
$U_F^1(0),\dots, U_F^1(3)$ are passed by $Z$, but $Z$ fails $U_F^1(4)$,
because after  execution trace
\[
s = a.c.c.c, \qquad\text{(note that $G(Z)/s = \text{Node\_3}$ according to Fig.~\ref{fig:tgZ})},
\]
it may be the case that $Z$ accepts only $\{b\}$ due to the internal choice
and $U_F^1(4)$ -- also due to internal choice -- accepts only the minimal
hitting set $\{ c \}$ and the event $a\in (\Sigma - [P/s]^0)$. So,
$(Z\parallel[\Sigma] U_F^1(4))/s$ deadlocks, and the $pass$ event cannot be
produced. Another failing execution arises if $Z/s$ chooses to accept only
$\{c \}$, while $U_F^1(4)/s$ choses to accept only $\{a,b\}$. Therefore,
\[
(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F^1(4)),
\]
and the test fails.
\xbox
\end{example}
\fixme{alcc: I would write down the actual test generated.}

### 2018-09-03 paper/conclusion.tex
small suggestions
added discussion + more suggestions

\section{Discussion}
\label{sec:conc}
% ==============================================================================
\subsubsection*{Discussion of Fault Domains}

@todo

% ==============================================================================
\subsubsection*{Discussion of Further Reductions of the Test Effort}
It is known from complete testing strategies for finite state machines that 
the upper bound $pq$ for the lengths of the 
traces used in our tests to investigate the SUT behaviour can be reduced. It is also
known from FSM testing that it is not necessary to test {\it all} traces up to this maximal length. Notable complete 
strategies supporting this fact have been presented, for example,
 in~\cite{hierons_testing_2004,DBLP:conf/forte/DorofeevaEY05,petrenko_testing_2011,simao_reducing_2012}. From~\cite{Huang2017} it
 is known that complete 
 FSM testing theories can be translated to other formalisms, such as 
Extended Finite State Machines, Kripke Structures, or CSP, resulting in likewise complete
test strategies for the latter. We intend to study translations of several promising FSM
strategies to CSP in the future. These will effectively reduce the upper bound 
$\ell\le |\Sigma|^k$ introduced in Section~\ref{sec:complexity}. The bound $h$ for
the number of sets to be used in probing the SUT for illegal deadlocks, however, cannot be further reduced, as has been established in Lemma~\ref{lemma:hseta}.

% ==============================================================================
\subsubsection*{Discussion of Adaptive Test Cases}
The tests suggested 
in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} were \emph{preset} in the
sense that the trace to be executed was pre-defined for each test. As a consequence,
the authors of~\cite{DBLP:conf/icfem/CavalcantiG07} introduced \emph{inconclusive}
as a third test result, applicable to the situations where the intended trace
of the execution was blocked, due to legal, but nondeterministic behaviour of the
SUT. In contrast to that, 
our test cases
specified in Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
adaptive. This has the advantage that test executions $(Q\parallel[\Sigma] U_F(p))$
for failures refinement
will never block before the final step specified by branch (\ref{eq:ufd}). As
a consequence, we do not need inconclusive test results. It should be noted, however,
that it is necessary for our test verdicts to recognise also deadlocks in the final
test step and interpret 
them as FAIL, as described in Section~\ref{sec:finitecompletefails}. In practice,
this is realised by adding a timeout event to the testing environment which indicates
deadlock situations. For real-time systems, this is an accepted technique, because
the SUT has to respond within a pre-defined latency interval, otherwise its behaviour 
is considered to be blocked and regarded as a failure. Our tests executions 
$(Q\parallel[\Sigma] U_T(p))$ for trace refinement never block at all. The adaptive       
behaviour of our test cases, however, induce the obligation to check that {\it all}
possible executions have been performed before the test can be considered as passed.
Typically, it is therefore assumed that a 
\emph{complete testing assumption}~\cite{hierons_testing_2004} holds which means that
every possible behaviour of the SUT has been performed after a finite number of
test executions. 
In practice, this is realised by executing each test several times, 
recording the traces that have been performed, and using hardware or software coverage 
analysers to determine whether all possible test execution behaviours of the SUT
have been observed. Therefore, the application of adaptive test cases comes at the price
of having to apply some grey-box testing techniques enabling us to decide whether
all SUT behaviours have been observed.

% ==============================================================================
\subsubsection*{Implications for CSP Model Checking}
As explained in the previous sections, passing a test is specified by
$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$
for failures testing. If the SUT $Q$ is not a programmed piece of software or
an integrated HW/SW system, but just another CSP process specification,
it is of course possible to verify these pass criteria using the FDR4 model checker.
For checking the refinement relation $P\lessdet_F Q$, the pass criteria have to be verified
for $k=0,\dots,pq-1$, where $p$ and $q$ indicate the number of nodes in $P$'s transition
graph and the maximal number of nodes in $Q$'s graph, respectively (Theorem~\ref{th:failurestest}). Since the test cases
$U_F(k)$ have such a simple structure, it is an interesting question for further 
research whether checking 
$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for $k=0,\dots,pq-1$ can be faster than directly 
checking $P\lessdet_F Q$, as one would do in the usual 
approach with FDR4. This is of particular interest, since the checks could be parallelised 
on several CPUs. Alternatively it is interesting to investigate whether the
the check\footnote{We are grateful to Bill Roscoe for suggesting this option.}
\[
(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] \Intchoice_{k=0}^{pq-1} U_F(k)) \hide \Sigma
\]
may perform better than the check of $P\lessdet_F Q$, since the former allows for 
other optimisations in the model checker than the latter.

@@

In this paper, we have introduced finite complete testing strategies for 
model-based testing against CSP reference models. The strategies are applicable   
to the conformance relations failures refinement and trace refinement. The underlying fault domains have been defined as the sets of CSP processes whose normalised transition
graphs did not have more than a given number of additional nodes, when compared to the transition graph of the reference process. For these domains, finite complete 
test suites are available. We have shown for the strategy checking failures refinement,
 that the way of probing the SUT for illegal deadlocks used in our test cases is 
 optimal in the sense that it is not possible to guarantee exhaustiveness with fewer
 probes.

@@

\paragraph{Acknowledgements}
The authors would like to thank Bill Roscoe and Thomas Gibson-Robinson for their advice on using the FDR4 model checker and for very helpful discussions concerning the potential implications of this paper in the field of model-checking.

>>>>

% ==============================================================================
\section{Discussion and Conclusions}
\label{sec:conc}
% ==============================================================================

% ==============================================================================
\subsubsection*{Discussion of Further Reductions of the Test Effort}
It is known from complete testing strategies for finite state machines that
the upper bound $pq$ for the lengths of the traces used in our tests to
investigate the SUT behaviour can be reduced. It is also known from FSM
testing that it is not necessary to test {\it all} traces up to this maximal
length. Notable complete strategies supporting this fact have been presented,
for example,
in~\cite{hierons_testing_2004,DBLP:conf/forte/DorofeevaEY05,petrenko_testing_2011,simao_reducing_2012}.
From~\cite{Huang2017} it is known that complete FSM testing theories can be
translated to other formalisms, such as Extended Finite State Machines,
Kripke Structures, or CSP, resulting in likewise complete test strategies for
the latter. We intend to study translations of several promising FSM
strategies to CSP in the future. These will effectively reduce the upper
bound $\ell\le |\Sigma|^k$ introduced in Section~\ref{sec:complexity}. The
bound $h$ for the number of sets to be used in probing the SUT for illegal
deadlocks, however, cannot be further reduced, as has been established in
Lemma~\ref{lemma:hseta}. \fixme{alcc: my guess is that these complete tests
are not for refinement. I think we can perhaps say something about this?}

% ==============================================================================
\subsubsection*{Discussion of Adaptive Test Cases}
The tests suggested
in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} were \emph{preset} in the
sense that the trace to be executed was pre-defined for each test. As a consequence,
the authors of~\cite{DBLP:conf/icfem/CavalcantiG07} introduced \emph{inconclusive}
as a third test result, applicable to the situations where the intended trace
of the execution was blocked, due to legal, but nondeterministic behaviour of the
SUT.

In contrast to that, our test cases specified in
Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
adaptive. This has the advantage that test executions $(Q\parallel[\Sigma]
U_F(p))$ for failures refinement never blocks before the final step specified
by branch (\ref{eq:ufd}), and so we do not need inconclusive test results. It
should be noted, however, that it is necessary for our test verdicts to
recognise also deadlocks in the final test step and interpret them as FAIL,
as described in Section~\ref{sec:finitecompletefails}. In practice, this is
realised by adding a timeout event to the testing environment which indicates
deadlock situations. For real-time systems, this is an accepted technique,
because the SUT has to respond within a pre-defined latency interval,
otherwise its behaviour is considered to be blocked and regarded as a
failure. \fixme{alcc: if deadlock is not acceptable?}

Our tests executions $(Q\parallel[\Sigma] U_T(p))$ for trace refinement never
block at all. The adaptive behaviour of our test cases, however, induce the
obligation to check that {\it all} possible executions have been performed
before the test can be considered as passed. Typically, it is therefore
assumed that a \emph{complete testing assumption}~\cite{hierons_testing_2004}
holds, which means that every possible behaviour of the SUT is performed
after a finite number of test executions. In practice, this is realised by
executing each test several times, recording the traces that have been
performed, and using hardware or software coverage analysers to determine
whether all possible test execution behaviours of the SUT have been observed.
Therefore, adaptive test cases come at the price of having to apply some
grey-box testing techniques enabling us to decide whether all SUT behaviours
have been observed.

% ==============================================================================
\subsubsection*{Discussion of Fault Domains}
As already mentioned, the work in~\cite{DBLP:conf/pts/CavalcantiS17} defines
a fault domain as the set of processes that refine a given CSP process.  In
that context, only testing for traces refinement is considered, and the
complete test suites may not be finite. So, the work presented here go well
beyond what is achieved there. On the other hand,
\cite{DBLP:conf/pts/CavalcantiS17} presents an algorithm for test generation
that can be easily adapted to consider additional selection and termination
criteria, like, for example, the length of the traces used to construct
tests. It would be possible, for instance, to use the bound indicated here.
Morevoer, specifying a fault domain as a CSP process allows us to model
domain-specific knowledge using CSP. For example, if an initialisation
component defined by a process $I$ can be regarded as correct without further
testing, we can use $I; RUN$ as a fault domain, to indicate that any SUT of
interest implements $I$ correctly, but afterwards has a potentially arbitrary
behaviour specified by $RUN$.

% ==============================================================================
\subsubsection*{Implications for CSP Model Checking}
As explained in the previous sections, passing a test is specified by
$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for
failures testing. If the SUT $Q$ is not a programmed piece of software or an
integrated hardware or software system, but just another CSP process
specification, it is of course possible to verify these pass criteria using
the FDR4 model checker. For checking the refinement relation $P\lessdet_F Q$,
the pass criteria have to be verified for $k=0,\dots,pq-1$, where $p$ and $q$
indicate the number of nodes in $P$'s transition graph and the maximal number
of nodes in $Q$'s graph, respectively (Theorem~\ref{th:failurestest}). Since
the test cases $U_F(k)$ have such a simple structure, it is an interesting
question for further research whether checking $(\epass\then\Stop) \lessdet_F
(Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for $k=0,\dots,pq-1$ can be faster
than directly checking $P\lessdet_F Q$, as one would do in the usual approach
with FDR4. This is of particular interest, since the checks could be
parallelised on several CPUs. Alternatively it is interesting to investigate
whether the check\footnote{We are grateful to Bill Roscoe for suggesting this
option.}
\[
(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] \Intchoice_{k=0}^{pq-1} U_F(k)) \hide \Sigma
\]
may perform better than the check of $P\lessdet_F Q$, since the former allows for
other optimisations in the model checker than the latter.

@@

In this paper, we have introduced finite complete testing strategies for
model-based testing against CSP reference models. The strategies are
applicable to the conformance relations failures refinement and trace
refinement. The underlying fault domains have been defined as the sets of CSP
processes whose normalised transition graphs do not have more than a given
number of additional nodes, when compared to the transition graph of the
reference process. For these domains, finite complete test suites are
available. We have shown for the strategy to check failures refinement that
the way of probing the SUT for illegal deadlocks used in our test cases is
optimal in the sense that it is not possible to guarantee exhaustiveness with
fewer probes.

@@

\paragraph{Acknowledgements}
The authors would like to thank Bill Roscoe and Thomas Gibson-Robinson for
their advice on using the FDR4 model checker and for very helpful discussions
concerning the potential implications of this paper in the field of model
checking. The work of Cavalcanti is funded by the Royal Academy of
Engineering and UK EPSRC Grant EP/R025134/1.

### 2018-09-03 paper/finitecomplete-failures.tex
added discussion + more suggestions

@@

We observe that the number of possible executions of $Q\parallel[\Sigma]
U_F(p)$ is finite, because the number of traces $s$ with maximal length
$(p+1)$ is finite and the sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and
$\minhits(P/s)$ are finite. We further recall that $\minhits(P/s)$ may be
empty, so that operand (\ref{eq:ufd}) of the external choice construct
defaults to $\Stop$. In this case, however, $\minaccs(P/s) = \{\varnothing\}$
applies, so the test executes the branch (\ref{eq:ufb}) and terminates with
$\epass$.

@@

After having run through a trace of length $p$, and if
$\minaccs(P/s)\neq\{\varnothing\}$ so that also
$\minhits(P/s)\neq\varnothing$, the test changes its behaviour: instead of
offering {\it all} legal events from $[P/s]^0$ to the SUT, it
nondeterministically chooses a minimal hitting set of $\minaccs(P/s)$ and
only offers the events contained in this set. If the SUT refuses to engage
into any of these events, this reveals a violation of the failures
refinement: according to Lemma~\ref{lemma:hseta}, a conforming SUT should
accept at least one event of each minimal hitting set in $\minhits(P/s)$.
Therefore, the test only terminates with  $\epass$, if such an event is
accepted by the SUT.

@@

A test suite $\TS$ is called \emph{complete} with respect to fault model ${\cal F}$,
if and only of the following conditions are fulfilled.
\begin{description}
\item[1.~Soundness] If $P \sqsubseteq Q$, then $Q$ passes all tests in $\TS$.
\item[2.~Exhaustiveness] If $P \not\sqsubseteq Q$ and $Q\in{\cal D}$,
then $Q$ fails at least one test in $\TS$.
\end{description}
%
In the next section, we establish the completeness of our test suite.

@@

\begin{theorem}\label{th:failurestest}
Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_F = \{ U_F(k)~|~0 \le k < pq  \}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
%
The proof of the theorem follows from the two lemmas below. The first states
that test suite $\TS_F$ is sound, the second states that the suite is also
exhaustive.
%
\begin{lemma}\label{lemma:mainfsound}
Test suite $\TS_F$ generated from CSP process $P$, as specified in
Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_F Q$.
\end{lemma}
\begin{proof}
{\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $P\ conf\
Q$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$,
any adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the external
choice construction for $U_F(p,s)$. Branch~(\ref{eq:ufa}) can never be
entered in the parallel execution of $Q$ and $U_F(p)$, because
$[Q/s]^0\subseteq [P/s]^0$ for all traces of $Q$.

>>>>

Here, we define our notion of tests for failures refinement, and then prove
completeness of our suite. Finally, we study to complexity of our approach by
identifying a bound on the number of tests we need in a complete suite.

@@

We observe that the number of possible executions of $Q\parallel[\Sigma]
U_F(p)$ is finite, because the number of traces $s$ with maximal length
$(p+1)$ is finite and the sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and
$\minhits(P/s)$ are finite. We further recall that $\minhits(P/s)$ may be
empty, so that operand (\ref{eq:ufd}) of the external choice construct
defaults to $\Stop$. In this case, however, $\minaccs(P/s) = \{\varnothing\}$
applies, so the test executes the branch (\ref{eq:ufb}) and terminates with
$\epass$. \fixme{alcc: Oops. This is undefined in CSP.}

@@

After having run through a trace of length $p$, and if
$\minaccs(P/s)\neq\{\varnothing\}$ so that also
$\minhits(P/s)\neq\varnothing$, the test changes its behaviour: instead of
offering {\it all} legal events from $[P/s]^0$ to the SUT, it
nondeterministically chooses a minimal hitting set of $\minaccs(P/s)$ and
only offers the events contained in this set. If the SUT refuses to engage
into any of these events, this reveals a violation of failures refinement:
according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least
one event of each minimal hitting set in $\minhits(P/s)$. Therefore, the test
only terminates with  $\epass$, if such an event is accepted by the SUT.

@@

A test suite $\TS$ is called \emph{complete} with respect to fault model ${\cal F}$,
if and only of the following conditions are fulfilled.
\begin{description}
\item[1.~Soundness] If $P \sqsubseteq Q$, then $Q$ passes all tests in $\TS$.
\item[2.~Exhaustiveness] If $P \not\sqsubseteq Q$ and $Q\in{\cal D}$,
then $Q$ fails at least one test in $\TS$.
\end{description}
%
In the next section, we establish the completeness of our test suite.
\fixme{alcc: I don't think we need a section just for this material.}

@@

\begin{theorem}\label{th:failurestest}
Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_F = \{ U_F(k)~|~0 \le k < pq  \}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
%
The proof of the theorem follows from the two lemmas below. The first states
that test suite $\TS_F$ is sound, the second states that the suite is also
exhaustive.
%
\begin{lemma}\label{lemma:mainfsound}
Test suite $\TS_F$ generated from a CSP process $P$, as specified in
Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_F Q$.
\end{lemma}
\begin{proof}
{\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $P\ conf\
Q$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$,
any adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the external
choice construction for $U_F(p,s)$. Branch~(\ref{eq:ufa}) can never be
entered in the parallel execution of $Q$ and $U_F(p)$, because
$[Q/s]^0\subseteq [P/s]^0$ for all traces of $Q$.
